
# Preprocessing
```{r}
full.data <- read.csv("./cleaned_data.csv", header = TRUE, sep = ",")
df <- full.data

############################
## drop compression rate
############################
df <- subset(df, select = -c(compression.ratio))

############################
## drop stroke
############################
df <- subset(df, select = -c(stroke))

############################
## Construct area form width
## and length. Then drop
## redundant values
############################
df$area <- df$width * df$length

df <- subset(df, select = -c(width, length))

############################
## Construct area form width
## and length. Then drop
## redundant values
############################
df$mpg <- rowMeans(df[, c("city.mpg", "highway.mpg")], na.rm = TRUE)

df <- df[, !colnames(df) %in% c("city.mpg", "highway.mpg")]

############################
## drop insurance  data
############################
# df <- subset(df, select = -c(normalized.losses, symboling))
df <- subset(df, select = -c(normalized.losses))


categorial_vars.all <- c("make", "symboling", "fuel.type", "aspiration",
                         "num.of.doors", "body.style", "drive.wheels", "engine.location",
                         "engine.type", "num.of.cylinders", "fuel.system")

for (cat_var in categorial_vars.all) {
  if (cat_var %in% colnames(df)) {
    df[[cat_var]] <- as.factor(df[[cat_var]])
  }
}


numerical_vars.all <- c("normalized.losses", "wheel.base", "length", "width", "height",
                        "area", "crub.weight", "engine.size", "bore", "stroke",
                        "horsepower", "peak.rpm", "city.mpg", "highway.mpg", "mpg", "price")

for (num_var in numerical_vars.all) {
  if (num_var %in% colnames(df)) {
    df[[num_var]] <- as.numeric(as.character(df[[num_var]]))
  }
}

# Summary statistics
summary(df)

# Check for missing values
colSums(is.na(df))

```

```{r}
library(FactoMineR)
library(factoextra)
library(NbClust)
library(mclust)
library(ggplot2)
library(cluster)
library(gridExtra)



numeric_df <- df[, sapply(df, is.numeric) & names(df) != "price"]
str(numeric_df)
scaled_numeric_df <- scale(numeric_df)

```


## HCPC
```{r}


# Perform PCA on numeric variables
res.pca <- PCA(numeric_df, ncp = 2, graph = FALSE)

res.hcpc <- HCPC(res.pca, graph = FALSE)

# Dendrogram using fviz_dend
fviz_dend(res.hcpc, rect = TRUE, rect_fill = TRUE, main="")

## 3D plot ##
# plot(res.hcpc, choice = "3D.map")

# Visualization of Clusters
fviz_cluster(res.hcpc,
             geom = "point",
             repel = FALSE, 
             show.clust.cent = TRUE, 
             main = "")  # Exclude text labels

```

```{r}
df$cluster <- res.hcpc$data.clust$clust

# Boxplot

boxplot_price <- ggplot(df, aes(x = factor(cluster), y = price, fill = cluster)) +
  geom_boxplot() +
  labs(title = "", x = "Cluster", y = "Price") +
  theme_minimal()

# Violin plot
violin_plot_price <- ggplot(df, aes(x = factor(cluster), y = price, fill = cluster)) +
  geom_violin() +
  labs(title = "", x = "Cluster", y = "Price") +
  theme_minimal()
print(boxplot_price)
print(violin_plot_price)
```


##  Hierarchical Clusters 
```{r}
# Transpose the data frame to compute distances between variables
# transposed_df <- t(scaled_numeric_df)

d <- dist(scaled_numeric_df)
dist_matrix_df <- as.data.frame(as.table(as.matrix(d)))

# Create a heatmap
ggplot(dist_matrix_df, aes(Var1, Var2, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Distance Between Points", x = "Observation", y = "Observation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
methods <- c("single", "complete", "average", "ward.D2", "centroid")
```

```{r}
# Initialize an empty list to store results
res_euclide.fit <- list()

# Loop through the methods and perform hierarchical clustering
for (method in methods) {
  # Compute hierarchical clustering
  fit <- hclust(d, method = method)
  
  # Add the clustering result to the list with the method name
  res_euclide.fit[[method]] <- fit
  
  # Plot the dendrogram
  plot(fit, main = paste("Dendrogram of", method, "Linkage"), xlab = "", sub = "")
}
 
```



### Ellbow selection of clusters
this was not helpful to devide the clusters
```{r}
for (method in names(res_euclide.fit)) {
  fit <- res_euclide.fit[[method]]  # Extract the hclust object
  
  # Plot the Elbow Method for the current fit
  plot(fit$height, 
       type = "b", 
       main = paste("Elbow Method for", method, "Linkage"), 
       # main ="",
       xlab = "Number of Clusters", 
       ylab = "Height")
}

```
#### NbClust
```{r}
library(NbClust)

# Determine the number of clusters
methods <- c("single", "complete", "average", "ward.D2", "centroid")
for (meth in methods) {
  nb <- NbClust(scaled_numeric_df, distance = "euclidean", min.nc = 2, max.nc = 10, method = meth)
  
}

```

```{r}
ks <- c(2, 2, 5, 2, 2)  # Number of clusters for each method

res_euclide.clusters <- list()

for (i in 1:length(methods)) {
  method <- methods[i]  # Get the current method
  fit <- res_euclide.fit[[method]]  # Retrieve the hclust object
  res_euclide.clusters[[method]] <- cutree(fit, k = ks[i])  # Cut the tree
}

```

```{r}
cluster_summaries <- list()

# Loop through clustering results
for (method in names(res_euclide.clusters)) {
  clusters <- res_euclide.clusters[[method]]  # Get cluster assignments
  
  # Add cluster assignments to the data
  df$cluster <- clusters
  
  # Summarize price statistics by cluster
  cluster_summary <- aggregate(price ~ cluster, data = df, 
                                FUN = function(x) c(mean = mean(x), median = median(x), sd = sd(x), n = length(x)))
  
  # Store the summary in the list
  cluster_summaries[[method]] <- cluster_summary
  print(cluster_summary)

}

```
### viz (Scatter and Boxplots)
```{r}

pca_coords <- as.data.frame(res.pca$ind$coord)

for (method in names(res_euclide.clusters)) {
  # Get cluster assignments
  clusters <- res_euclide.clusters[[method]]
  
  # Add cluster assignments to the data
  df$cluster <- as.factor(clusters)
  
  # Create a boxplot of price by cluster
  p <- ggplot(df, aes(x = cluster, y = price, fill = cluster)) +
    geom_boxplot() +
    labs(title = paste("Price Distribution by Clusters (Method:", method, ")"),
         x = "Cluster", y = "Price") +
    theme_minimal() +
    theme(legend.position = "none")
  
  # Print the plot explicitly
  print(p)
  pca_coords[[method]] <- as.factor(clusters)  # Store clusters for each method
  
  p <- ggplot(pca_coords, aes(x = Dim.1, y = Dim.2, color = pca_coords[[method]])) +
    geom_point(size = 3, alpha = 0.7) +
    labs(title = paste("Clusters Visualized on PCA Dimensions (Method:", method, ")"),
         x = "PCA Dimension 1", y = "PCA Dimension 2", color = "Cluster") +
    theme_minimal()
    
  # Print the plot explicitly
  print(p)
  
    # Create a bar plot for cluster sizes
  cluster_sizes <- as.data.frame(table(clusters))
  print(cluster_sizes)
  
  colnames(cluster_sizes) <- c("Cluster", "Count")
  
  p <- ggplot(cluster_sizes, aes(x = Cluster, y = Count, fill = Cluster)) +
    geom_bar(stat = "identity") +
    labs(title = paste("Cluster Sizes (Method:", method, ")"),
         x = "Cluster", y = "Count") +
    theme_minimal() +
    theme(legend.position = "none")
  
  # Print the bar plot
  print(p)
}

######### Clustering of hc clustering

# Extract HCPC cluster assignments
df$cluster <- as.factor(res.hcpc$data.clust$clust)

# Boxplot of price by cluster
p <- ggplot(df, aes(x = cluster, y = price, fill = cluster)) +
  geom_boxplot() +
  labs(title = "Price Distribution by Clusters (HCPC Method)",
       x = "Cluster", y = "Price") +
  theme_minimal() +
  theme(legend.position = "none")
print(p)

# Add HCPC clusters to PCA coordinates
pca_coords$HCPC_cluster <- as.factor(res.hcpc$data.clust$clust)

# Scatter plot of clusters on PCA dimensions
p <- ggplot(pca_coords, aes(x = Dim.1, y = Dim.2, color = HCPC_cluster)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "Clusters Visualized on PCA Dimensions (HCPC Method)",
       x = "PCA Dimension 1", y = "PCA Dimension 2", color = "Cluster") +
  theme_minimal()
print(p)

# Bar plot of cluster sizes
cluster_sizes <- as.data.frame(table(res.hcpc$data.clust$clust))
colnames(cluster_sizes) <- c("Cluster", "Count")

p <- ggplot(cluster_sizes, aes(x = Cluster, y = Count, fill = Cluster)) +
  geom_bar(stat = "identity") +
  labs(title = "Cluster Sizes (HCPC Method)",
       x = "Cluster", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none")
print(p)



```
### Plots without title for report
```{r}
for (method in c("average", "ward.D2")){
  clusters <- res_euclide.clusters[[method]]
  
  # Add cluster assignments to the data
  df$cluster <- as.factor(clusters)
  
  # Create a boxplot of price by cluster
  p <- ggplot(df, aes(x = cluster, y = price, fill = cluster)) +
    geom_boxplot() +
    labs(title = "",
         x = "Cluster", y = "Price") +
    theme_minimal() +
    theme(legend.position = "none")
  
  # Print the plot explicitly
  print(p)
  pca_coords[[method]] <- as.factor(clusters)  # Store clusters for each method
  
  p <- ggplot(pca_coords, aes(x = Dim.1, y = Dim.2, color = pca_coords[[method]])) +
    geom_point(size = 3, alpha = 0.7) +
    labs(title = "",
         x = "PCA Dimension 1", y = "PCA Dimension 2", color = "Cluster") +
    theme_minimal()
    
  # Print the plot explicitly
  print(p)
}

```


## K means

```{r}
head(df)
```


```{r}
# Perform k-means clustering
set.seed(123) # For reproducibility
kmeans_result <- kmeans(scaled_numeric_df, centers = 3 , nstart = 25)

# Add cluster assignments to the original dataframe
df$cluster <- as.factor(kmeans_result$cluster)


cluster_sizes <- as.data.frame(table(kmeans_result$cluster))
print(cluster_sizes)
```


```{r}
# Bar plot of cluster sizes
cluster_sizes <- as.data.frame(table(df$cluster))
colnames(cluster_sizes) <- c("Cluster", "Count")

p <- ggplot(cluster_sizes, aes(x = Cluster, y = Count, fill = Cluster)) +
  geom_bar(stat = "identity") +
  labs(title = "",
       x = "Cluster", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none")
print(p)

# Boxplot of price by cluster
p <- ggplot(df, aes(x = cluster, y = price, fill = cluster)) +
  geom_boxplot() +
  labs(title = "",
       x = "Cluster", y = "Price") +
  theme_minimal() +
  theme(legend.position = "none")
print(p)

# Add cluster assignments to PCA coordinates
pca_coords$kmeans <- df$cluster

# Scatter plot of clusters on PCA dimensions
p <- ggplot(pca_coords, aes(x = Dim.1, y = Dim.2, color = kmeans)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "",
       x = "PCA Dimension 1", y = "PCA Dimension 2", color = "Cluster") +
  theme_minimal()
print(p)

  

```


```{r}
# Save clusters 

df$kmeans <- as.factor(kmeans_result$cluster)
df$hcpc <- as.factor(res.hcpc$data.clust$clust)

clusters <- res_euclide.clusters[["complete"]]
# Add cluster assignments to the data
df$hc.complete <- as.factor(clusters)

clusters <- res_euclide.clusters[["average"]]
# Add cluster assignments to the data
df$hc.avg <- as.factor(clusters)

clusters <- res_euclide.clusters[["ward.D2"]]
# Add cluster assignments to the data
df$hc.ward.D2 <- as.factor(clusters)
```

```{r}
str(df)
write.csv(df, file = "./clustered_data.csv", row.names = FALSE)
```



## Cluster Evaluation
### Sillouette
```{r}

# List of cluster variables
cluster_vars <- c("kmeans", "hcpc", "hc.complete", "hc.avg", "hc.ward.D2")

# Initialize a data frame to store silhouette scores
silhouette_results <- data.frame(Cluster_Method = character(), Silhouette_Score = numeric())

# Compute silhouette scores for each clustering method
for (cluster_var in cluster_vars) {
  # Calculate silhouette scores
  silhouette_scores <- silhouette(as.numeric(df[[cluster_var]]), dist(scaled_numeric_df))
  
  # Compute the mean silhouette score
  avg_silhouette <- mean(silhouette_scores[, "sil_width"])
  
  # Store the result
  silhouette_results <- rbind(silhouette_results, data.frame(
    Cluster_Method = cluster_var,
    Silhouette_Score = avg_silhouette
  ))
}

# Bar plot of silhouette scores
p <- ggplot(silhouette_results, aes(x = Cluster_Method, y = Silhouette_Score, fill = Cluster_Method)) +
  geom_bar(stat = "identity") +
  labs(title = "Silhouette Scores for Different Clustering Methods",
       x = "Clustering Method", y = "Average Silhouette Score") +
  theme_minimal() +
  theme(legend.position = "none")
print(p)
print(silhouette_results)

```

```{r}

# List of cluster variables
cluster_vars <- c("kmeans", "hcpc", "hc.complete", "hc.avg", "hc.ward.D2")

# Initialize a data frame to store ARI scores
ari_results <- data.frame(Cluster_Method = character(), ARI_Score = numeric())

# Compute ARI for each clustering method
for (cluster_var in cluster_vars) {
  # Calculate ARI
  ari_score <- adjustedRandIndex(df[[cluster_var]], df$symboling)
  
  # Store the result
  ari_results <- rbind(ari_results, data.frame(
    Cluster_Method = cluster_var,
    ARI_Score = ari_score
  ))
}

# Bar plot of ARI scores
p <- ggplot(ari_results, aes(x = Cluster_Method, y = ARI_Score, fill = Cluster_Method)) +
  geom_bar(stat = "identity") +
  labs(title = "Adjusted Rand Index (ARI) for Different Clustering Methods",
       x = "Clustering Method", y = "ARI Score") +
  theme_minimal() +
  theme(legend.position = "none")
print(p)
print(ari_results)
```

### Chi Squared test of independence for price
```{r}
# Create an empty data frame to store the results
chi_square_results <- data.frame(
  Clustering_Method = character(),
  P_Value = numeric(),
  stringsAsFactors = FALSE
)

# List of clustering methods
clustering_methods <- c("kmeans", "hcpc", "hc.complete", "hc.avg", "hc.ward.D2")

# Perform Chi-Square Test of Independence for each clustering method
for (method in clustering_methods) {
  # Perform Chi-Square test between price category and cluster assignment
  chi_result <- chisq.test(table(df$price, df[[method]]))
  
  # Store the results in the dataframe
  chi_square_results <- rbind(chi_square_results, 
                              data.frame(Clustering_Method = method, P_Value = chi_result$p.value))
}

# Print the table of p-values
print(chi_square_results)

```

### Kruskal Wallis test

```{r}
# Create an empty data frame to store the results
kruskal_results <- data.frame(
  Clustering_Method = character(),
  P_Value = numeric(),
  stringsAsFactors = FALSE
)

# Perform Kruskal-Wallis Test for each clustering method
for (method in clustering_methods) {
  # Perform Kruskal-Wallis test between price and cluster assignment
  kruskal_result <- kruskal.test(price ~ df[[method]], data = df)
  
  # Store the results in the dataframe
  kruskal_results <- rbind(kruskal_results, 
                           data.frame(Clustering_Method = method, P_Value = kruskal_result$p.value))
}

# Print the table of p-values
print(kruskal_results)


```


## Symboling Cluster Analysis

```{r}
cluster_vars <- c("kmeans", "hcpc", "hc.complete", "hc.avg", "hc.ward.D2")

# Visualize contingency matrices as bar plots
for (cluster_var in cluster_vars) {
  # Create a data frame for plotting
  contingency_table <- table(Cluster = df[[cluster_var]], Symboling = df$symboling)

  contingency_df <- as.data.frame(table(Cluster = df[[cluster_var]], Symboling = df$symboling))
  
  # Bar plot
  p <- ggplot(contingency_df, aes(x = Symboling, y = Freq, fill = Cluster)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Contingency Visualization:", cluster_var),
         x = "Symboling", y = "Frequency", fill = "Cluster") +
    theme_minimal()
  
    # Heatmap
  p <- ggplot(contingency_df, aes(x = Symboling, y = Cluster, fill = Freq)) +
    geom_tile() +
    scale_fill_gradient(low = "white", high = "blue") +
    labs(title = paste("Contingency Heatmap:", cluster_var),
         x = "Symboling", y = "Cluster", fill = "Frequency") +
    theme_minimal()
  
  # Compute relative frequencies (row-wise proportions)
  relative_table <- prop.table(contingency_table, margin = 1)
  
  # Convert to a data frame for plotting
  relative_df <- as.data.frame(as.table(relative_table))
  
  
    # Heatmap
  p <- ggplot(relative_df, aes(x = Symboling, y = Cluster, fill = Freq)) +
    geom_tile() +
    scale_fill_gradient(low = "white", high = "blue") +
    labs(title = paste("Contingency Heatmap:", cluster_var),
         x = "Symboling", y = "Cluster", fill = "Frequency") +
    theme_minimal()

  
  # Print the plot
  print(p)
  
  
# Boxplot of price by cluster
p <- ggplot(df, aes(x = get(cluster_var), y = price, fill = get(cluster_var))) +
  geom_boxplot() +
  labs(title = "",
       x = "Cluster", y = "Price") +
  theme_minimal() +
  theme(legend.position = "none")
print(p)
}

```


```{r}
cluster_vars <- c("kmeans", "hcpc", "hc.complete", "hc.avg", "hc.ward.D2")

# Visualize contingency matrices as bar plots
for (cluster_var in cluster_vars) {

p <- ggplot(df, aes(x = get(cluster_var), y = price, fill = cluster)) +
  geom_violin(trim = FALSE) +
  labs(title = "Density of Variable by Cluster",
       x = "Cluster", y = "Variable") +
  theme_minimal()
print(p)


```





TOOD: 

hypotheses test for clustering and price

printout a csv inclduing the clusterings so that riccardo can work on the same clusters as I did.



























